---
title: "Ring (HKCERT CTF 2022) Write-up, for O0031 (AUTOEXEC.BAT)"
output: md_document
date: "2022-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ring

> Ring ring ring ring ring ring ring banana phone
> - @ozetta

## Question

This is the entirety of `Ring.R`:

```
Ring <- function(p){
  tryCatch(
  expr = {
    x <- 1:17
    y <- utf8ToInt(p)
    z <- lm(y ~ poly(x,16))
    u <- as.numeric(summary(z)$coefficients[,1])
    v <- c(94.8823529411764923, -8.0697024752598967, -8.6432639293214333, 
            3.8067684547541667, -2.6157531995857521, 39.7193457764808500, 14.9176635631982180, 
           14.3308668599120725, 43.6042210530751291, 37.5259918448356302, 
            4.0314998333763086,  5.1052914400636569,  1.8689828029874489, 13.7270919105349307, 
           12.8538529135203099,  7.1197700159123247, 10.2656771598556720)
    if(identical(u, v)){
      print(sprintf("hkcert22{%s}", sprintf(p, "ReveRseengineeRing")))
    }else{
      stop()
    }
  },
  error = function(e){
    print("Wrong Password")
  },
  warning = function(w){
    print("Wrong Password")
  })
}
Ring(readline(prompt="Enter Password: "))
```

So, we basically have to reverse-engineer a set of 17 characters in the
printable ASCII range, so that when running a LM upon, we would get the given
coefficient `v`.

## Solution

Literally just two weeks ago (really!), I learned from a stats course about
`optim` and `optimize` functions in R, which sounds like the perfect approach.

We used the [SSE](https://en.wikipedia.org/wiki/Residual_sum_of_squares) to
measure how close our answer is to the flag.

For initial values, we chose 85, because it is a midpoint between 48 (`0`) and
122 (`z`), which we had confidence that most of the letters of the flag are
nearby.

Let's give it a quick shot:

```{r}
fn <- function (y) {
  x <- 1:17
  z <- lm(y ~ poly(x,16))
  u <- as.numeric(summary(z)$coefficients[,1])
  v <- c(94.8823529411764923, -8.0697024752598967, -8.6432639293214333, 
         3.8067684547541667, -2.6157531995857521, 39.7193457764808500, 14.9176635631982180, 
         14.3308668599120725, 43.6042210530751291, 37.5259918448356302, 
         4.0314998333763086,  5.1052914400636569,  1.8689828029874489, 13.7270919105349307, 
         12.8538529135203099,  7.1197700159123247, 10.2656771598556720)
  sum((u-v) ** 2)
}

optim_fn <- function (init) {
  res <- optim(init, fn, control = c(maxit = 10000, abstol = 0.1))
  res$par
}

score_par <- function (par) {
  print(par)
  print(fn(par))
  print(intToUtf8(round(par)))
}
```

```{r}
par1 <- optim_fn(rep(c(85), 17))
score_par(par1)
```

Looks promising, but not close enough. So let's re-run it with rounded values,
since we know the source strings are represented in integers anyways.

```{r}
par2 <- optim_fn(round(par1))
score_par(par2)
```

```{r}
par3 <- optim_fn(round(par2))
score_par(par3)
```

```{r}
par4 <- optim_fn(round(par3))
score_par(par4)
```

```{r}
par5 <- optim_fn(round(par4))
score_par(par5)
```

At this point here a few iterations down the road, we also tried to impose a
penalty to encourage the optimizer to give us values closer to integer values.

```{r}
fn_penalized <- function (y) {
  x <- 1:17
  z <- lm(y ~ poly(x,16))
  u <- as.numeric(summary(z)$coefficients[,1])
  v <- c(94.8823529411764923, -8.0697024752598967, -8.6432639293214333, 
         3.8067684547541667, -2.6157531995857521, 39.7193457764808500, 14.9176635631982180, 
         14.3308668599120725, 43.6042210530751291, 37.5259918448356302, 
         4.0314998333763086,  5.1052914400636569,  1.8689828029874489, 13.7270919105349307, 
         12.8538529135203099,  7.1197700159123247, 10.2656771598556720)
  sum((u-v) ** 2) + (sum((y - round(y) ** 2)) * 1e2)
}

optim_fn_penalized <- function (init) {
  res <- optim(init, fn_penalized, control = c(maxit = 10000, abstol = 0.1))
  res$par
}
```

Also, from the original question, we believed that there should be a `%s`
somewhere in the string that we are finding. Since `%` is 37 and `s` is 115, we
think it is most probable for those to be the 16th and 17th characters.

Let's try again.

```{r}
round_with_fixed_suffix <- function (par) {
  par[16] <- 37  # `%`
  par[17] <- 115 # `s`
  round(par)
}
par6 <- optim_fn_penalized(round_with_fixed_suffix(par5))
score_par(par6) # the score pumped because we are scoring with `fn`
```
We also tried alternating between the penalized function and the unpenalized
functions to hopefully finding a value that converges towards the flag.

```{r}
par7 <- optim_fn(round_with_fixed_suffix(par6))
score_par(par7)
```

```{r}
par8 <- optim_fn_penalized(round_with_fixed_suffix(par7))
score_par(par8)
```
This looks super promising, since there is a `%s` at the end after running it.
Let's try to optimize with `fn` once more:

```{r}
par9 <- optim_fn(round_with_fixed_suffix(par8))
score_par(par9)
```
Aha! Just to confirm:

```{r}
score_par(round(par9))
```
The numbers are not being completely aligned, but we knew we found our
jackpot. Let's try the flag out:

```{r}
p <- 'RegreSSSionMean%s'
sprintf("hkcert22{%s}", sprintf(p, "ReveRseengineeRing"))
```

Yay!

Gotta love the upcoming stats courses a bit more. But before that, back to my
weekly lab assignments (ergh).
